Overview and Summary of the First Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE1)

### 2014-03-28 WSSSPE1_report/WSSSPE1_summary_report.tex
Just some quick summaries of papers in prep for more systematic summary

\note{Question of extent to which this section (indeed the whole document) should call out to existing work on these questions, or stick to summarizing the papers and the discussion. I think we should call out as much as possible --JamesHowison}

\note{Dan: I think this is a report on the workshop, and as we need to refer to other things, that's fine, but the focus should be on what was submitted and discussed.}

@@

@@

@@

>>>>

@@

They examine existing ISO standards about software quality.  These standards include attributes of the software but currently do not discuss sustainability, in terms of both meanings of sustainability (energy efficiency and what they call perduability, the software "lasting over time". In terms of lasting over time, they highlight characteristics of reliabiity, maintainability (e.g., modularity, the degree to which the code is understandable, testable and modifiable) and the degree to which the code can be adapted.

@@

They argue that the way that a project is run relates to whether it can continue to serve its users. In particular they argue that because sustainability relates to having ongoing resources, governance must be open to receive diverse input and thus have the potential to "transform passive users into active contributors".

They also argue that open source principles have been mis-applied by the research software community.

Their framework puts together three parts supported by open software: operations and support, escience research and scientific collaboration.

They define open governance as "project deliberations on open, archivable forums ... with votes carried out asynchronously over a period of time ..."  This, presumably, contrasts with how governance is done in research software projects at present, through offline meetings amongst grant funded groups, for the very large part.

@@

Their framework looks at (for each software element): technical (how does it work with other components), social (who is building it, who is using it) and political context (who is funding it).

The propose three axes on which to locate particular pieces of software: temporal extent over which software should persist, spatial extent (particular lab to global context), Purpose (particular use to general use.).  They also discuss resourcing models: 1) commercial for profit, 2) open source, 3) closed partnerships and 4) funded projects.  An additional model can function during a sustaining phase: open source with paid support. Finally they introduce two models of governance, top-down and bottom-up.

They hypothesize that each of these models will be appropriate for a different location on their axes and look to additional research to make the connections (for example, government funding might be most appropriate for longer time scales).
